Multiway Spherical Clustering via \\ Degree-Corrected Tensor Block Models

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/response_s1.tex

%\begin{example}[Failure of  $\Delta_{\min}^2$ with unbalanced degree] Consider the order-2 dTBM of 2 communities with core matrix
%\begin{equation}
%    \mS = \begin{pmatrix} 1 & a\\
%    1 & -a
%    \end{pmatrix}, \quad \text{and} \quad \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = M \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
%\end{equation}
%where $\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}. Let $\Delta_{\mX}^2$ denote the angle between the mean tensor, where 
%\begin{equation}\label{eq:delta_x}
%    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \mX = \mat(\tX). %\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
%\end{equation}
%With core matrix $\mS$, we have 
%\begin{equation}
%    \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2}, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{ \onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2}.
%\end{equation}
%Take $ a = p^{-1/4}, M = p^{1/2 + \epsilon}$ for some $\epsilon > 0$. We have \begin{equation}
%    \Delta_{\min}^2 = p^{-1/2} \quad \text{and} \quad    \Delta_{\mX}^2 = p^{-1-\epsilon}.
%\end{equation}
% \begin{align}
%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 =  \Delta_{\mX}^2 = p^{-1}; \\
%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \Delta_{\mX}^2 = p^{-1}.
% \end{align}
%By following Theorem~\ref{thm:stats}, no estimator exact recovers dTBM under $\Delta_{\mX}^2 < p^{-1}$. However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
%\end{example}

%\begin{rmk}[Flexibility in balanced degree] One important note is that the balance assumption~\eqref{eq:degree} does not preclude the dramatic degree heterogeneity. In fact, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$. This range constraints on $\mtheta$ is looser than previous work \citep{gao2018community} that restricts the highest degree in a sub-linear rate $o(p)$. 
%\end{rmk}

%\begin{rmk}[Similar assumption in literature]
%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
%Similar regulations are not rare in literature. In higher-order model \cite{ke2019community}, similar assumption ${\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}\leq C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}$ is also made to ensure the degree distributions are balanced on each community. Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees. "
%\end{rmk}

%\end{quote}

New Section~\ref{sec:statlimit}:

\begin{quote}
``... We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP(\gamma)$, where 
\begin{equation}\label{eq:mle}
    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP (\gamma)} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
\end{equation}

\begin{customthm}{2}[Statistical critical value] Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition.

\begin{itemize}[wide]
    \item \textbf{Impossibility.} 
%     Considerthe minimal gap between the mean tensor as 
% \begin{equation}
%     \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \text{where} \quad \mX = \mat(\tX).
% \end{equation}
Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
\begin{equation}
  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\lambda)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
\item \textbf{MLE achievability.} If the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$, then, the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
\begin{equation}
\gamma> -(K-1)+c \quad  \Rightarrow  \quad \ell(\hat z_{\text{MLE}}, z) \lesssim \text{SNR}^{-1}\exp\left(-{p^{K-1}\text{SNR}\over r^{K-1}}\right) \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{customthm}

The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery when $\gamma < -(K-1)$. Inspection of the proof shows that the impossibility holds with or without the balanced degree assumption. %Moreover, with general degree, Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$. 
The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.

The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c$ for an arbitrary constant $c>0$. Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR. "
\end{quote}

New Section~\ref{sec:thm5}:
\begin{quote}
``...The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}. we first show a weaker polynomial rate for MLE, and then improve the rate from polynomial to exponential throughout the iterations. The only difference is that the MLE remains the same over iterations due to its global optimality. See Appendix II, Section~\ref{sec:statprove2} for the detailed proof.''
\end{quote}

New Section~\ref{sec:complimit}:

\begin{quote}
``...
\begin{customthm}{3}[Computational critical value] Consider general Gaussian dTBMs under the parameter space $\tP$ with $K\geq 2$. Then, we have the following computational phase transition.

\begin{itemize}[wide]
    \item \textbf{Impossibility.} Assume HPC conjecture holds. If the signal exponent satisfies $\gamma < -K/2$, then, no \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ achieves exact recovery in expectation as $p\to \infty$; that is, 
\begin{align}
  \gamma<-K/2 \ \Rightarrow \ \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
\end{align}
\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c$ for an arbitrary constant $c>0$. Furthermore, assume the degree is lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and $\mtheta$ satisfies the local linear stability in Definition~\ref{def:stable} in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. Then, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
\begin{equation}
     % \ell(\hat z_{\text{poly}}, z) \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 }.
  \gamma>-K/2+c\ \Rightarrow   \  \ell(\hat z_{\text{poly}}, z) \lesssim \text{SNR}^{-1}\exp \of{ - \frac{p^{K-1}\text{SNR}}{r^{K-1}} } \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{customthm}

The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when $\gamma > -K/2+c$ under extra technical assumptions. In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting. Therefore, we conclude that $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM with respect to our SNR. ''
\end{quote}

\end{enumerate}

\newpage

\begin{center}
    \textbf{Point-by-point Response to Reviewer 1}
\end{center}

\textit{The paper studies the problem of finding the planted block structure in a tensor, generated from the degree-corrected tensor block model (dTBM). This problem arises in different contexts, such as community detection in hypergraph, higher-order clustering etc. A simpler version of the problem (non-degree corrected TBM) has been studied in previous works. The main contribution of the paper is to extend the guarantees for recovering planted blocks in TBM to the case of dTBM. In particular, a suitable SNR is proposed and statistical lower bound (Thm 2) and computational lower bound under HPC (Thm 3) are derived. Thm 4-5 provide error rate for a poly-time algorithm.}

\textit{Overall, the paper is well-written and comes with additional numerical studies, also on real data sets. The contribution of the paper is quite good. There are few aspects that are not addressed in the paper (listed below). It is to be noted that all the mentioned may not be easily resolved, but brief discussions on these aspects would improve the paper.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]

\item \textit{The statistical upper bound (guarantees for MLE) seem to be missing since Thm 4-5 focuses on a poly-time approach. Please explicitly describe if (and how) one can derived error rate for MLE from Thm 5.}

\textbf{Response:} Thank you for the suggestion. We have added the MLE guarantee in the revised Theorem~\ref{thm:stats}, Section~\ref{sec:statlimit}. Detailed proof has been added in the Appendix II, Section~\ref{sec:statprove2}. The proof shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}; the only difference is that the MLE remains the same over iterations due to its global optimality. See our response to \textbf{\# 2 Major Comments from Editor} for details.

%The MLE guarantee matches with the statistical lower bound, and thus rigorously supports our claim that the statistical critical value is $\gamma_{\text{stat}} = -(K-1)$.

\item \textit{While the dTBM model is general, it ignores the effect of sparsity in the context of hypergraphs. In case of community detection in sparse hypergraphs, $\bbE[\tY]$ includes a scalar that vanishes with number of nodes/tensor dimension. As in the case of graph SBM, the analysis gets a bit tricky for sparse hypergraphs (which is not covered in this paper). The authors could look at \cite{florescu2016spectral} which studies this setting and the proposed method also has similarities with the algorithm presented here.}

\textbf{Response:} We thank reviewer for the suggestions. Our current work focuses on \emph{dense} dTBM. While sparse dTBM is an interesting application, the theory and algorithm require different techniques. 
Inspired by the comment, we have extended our algorithm and experiment to sparse dTBM with a decaying sparse parameter. We adapt diagonal-deleted HOSVD as a new initialization strategy under extremely sparse settings.

We have added the following discussion in Section \ref{subsec:ber}:

\begin{quote}
    ``... \textit{Extension to sparse binary dTBMs.} The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community}. Specifically, the sparse binary dTBM assumes that, the entries of $\tY$ follow independent Bernoulli distributions with the mean
     \begin{equation}\label{eq:sparse_dtbm}
    \bbE\tY = \alpha_p \tS \times_1 \mTheta \mM \times_2  \cdots \times_K \mTheta \mM,
\end{equation}
where the extra scalar parameter $\alpha_p \in (0,1]$ is function of $p$ that controls the sparsity. A smaller $\alpha_p$ indicates a higher level of sparsity. Our current work focuses on dense dTBM with $\alpha_p=1$.  While sparse dTBM is an interesting application, the algorithm and its analysis require different techniques. Below, we discuss possible modifications of the algorithm.

%The parameters $(z, \mtheta, \alpha_p \tS)$ are still identifiable up to permutation under our Assumption~\ref{assmp:min_gap}. However, the constant lower bound assumption on the core tensor is no longer suitable to describe the decaying sparse parameter. Therefore, our phase transition will not directly apply; we expect the extra sparsity parameter $\alpha_p$ should also be included in the signal notion, in order to characterize the phase transition.

The sparsity affects our initialization guarantee in our Theorem~\ref{thm:initial}. In our initialization, the spectral denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) implements matrix SVD to unfolded tensors. %, and Theorem~\ref{thm:initial} indicates the good performance under the dense case. 
However, SVD-based methods are believed to fail in extremely sparse SBM due to the localization phenomenon in the singular vectors \citep{florescu2016spectral}. Inspired by \cite{florescu2016spectral}, we adapt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.

The sparsity also affects the iteration guarantee in our Theorem~\ref{thm:refinement}. The decaying mean tensor leads to a worse statistical error of order $\tO( -\alpha_p p^{K-1})$ on $\hat \tX$. The theoretical analyses for sparse binary dTBM and algorithms are left as future direction. Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix I."
\end{quote}

We have added a numerical experiment and Figure~\ref{fig:sparse} under the sparse binary dTBM in Appendix I:

\begin{quote}

\begin{figure}[h!]
    \centering
    \includegraphics[width=.8\columnwidth]{sparsity.pdf}
    \caption{CER comparison versus sparsity parameter $\alpha_p$ in $[0.05, 0.9]$. We set $p = 100, r = 5$, and $\gamma = -1.2$ under sparse binary dTBM.}
    \label{fig:sparse}
\end{figure}

`` ...The second additional experiment evaluates the algorithm performances under the sparse binary dTBM~\eqref{eq:sparse_dtbm}. We fix the signal exponent $\gamma = -1.2$ and vary the sparsity parameter $\alpha_p \in [0.05, 0.9]$. A smaller $\alpha_p$ leads to a higher probability of zero entries in the observation. In addition to the three algorithms mentioned in Section~\ref{subsec:comp}  (denoted {\bf \small Initialization}, {\bf \small dTBM}, and {\bf \small SCORE}), we consider other three algorithms based on the discussion in Section~\ref{subsec:ber}: 
  \begin{itemize} 
  \item \textbf{\small D-HOSVD}, the diagonal-deleted HOSVD in \cite{ke2019community}; 
  \item \textbf{\small D-HOSVD + Angle}, the combined algorithm of our angle-based iteration with initialization from \textbf{\small D-HOSVD};
  \item \textbf{\small SCORE + Angle}, the combined algorithms of our angle-based iteration with initialization from \textbf{\small SCORE}.
  \end{itemize}

Figure~\ref{fig:sparse} shows a slightly larger error in \textbf{\small dTBM} than that in \textbf{\small SCORE}, \textbf{\small D-HOSVD + Angle}, and \textbf{\small SCORE + Angle} under the sparse setting with $\alpha_p < 0.3$. 
%The outperformance of \textbf{\small D-HOSVD} based iterative methods confirms the benefits of diagonal deletion in the sparse case. Meanwhile, 
The small gap between  \textbf{\small dTBM} and other sparse-specific methods implies the robustness of our algorithm. In addition, comparing \textbf{\small SCORE} versus \textbf{\small SCORE + Angle} (or \textbf{\small D-HOSVD} versus \textbf{\small D-HOSVD + Angle}) indicates the benefit of our angle iterations under the sparse dTBM. In the intermediate and dense cases with $\alpha_p \geq 0.3$, our proposed \textbf{\small dTBM} has a clear improvement over others, which again verifies the success of our algorithm in dense settings."
\end{quote}

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/response_s1.tex

\item \textit{ (Critical) The statistical lower bound is derived under non-degree TBM (since all $\theta(i) = 1$). From this, it is still not clear whether the lower bound would signiﬁcantly increase when there is a skewed degree correction. The authors’ claim that the statistical lower bound matches MLE is diﬃcult to verify (since I could not ﬁnd an explicit result for MLE). Moreover, the stated upper bound (Thm 4-5) assume some uniformity in $\theta$ (see top of page-11), and so the (matched) upper/lower bounds could only be for the case of relatively homogeneous degrees. A more clear discussion on this would be useful. The present results leaves open the possibility that phase-transition SNR for dTBM could be arbitrarily worse than that of TBM, in which case making some assumption on balance is reasonable.}

\textbf{Response:}

% \fixme{Miaoyan}{Say ``There seem to be three questions in this comments. We address them one by one. 
% \begin{enumerate}
% \item Regarding ``whether the lower bound would significatly increase when there is a skewed degree correction''. (move your response here)
% \item Regarding ``explicit result for MLE'' (move your response here)
% \item Regarding ``The stated upper bound assume some...''. 
% \end{enumerate}}

We thank reviewer for the suggestions. There seem to be three questions in this comment. We address them one by one.

\begin{enumerate}
    \item Regarding ``whether the lower bound would significantly increase when there is a skewed degree correction". We have revised the proof of statistical minimax bound with relaxed degree assumptions. Briefly, we relax the previous non-degree construction $\theta(i)\equiv 1$ to the skewed-degree construction under assumption~\eqref{eq:degree}. The latter allows moderate degree heterogeneity; for example, within each of the cluster, we allow the highest degree to be $\theta(i) = \Omega(p)$, whereas the lowest degree to be $\theta(i)=\tO(1)$. The revised minimax bound does not increase in the presence of moderately skewed degree. See our new Theorem~\ref{thm:stats} in Section~\ref{sec:limits} and new proof in Appendix II, Section~\ref{sec:statprove2}.

    \item  Regarding ``explicit result for MLE". We have added the MLE guarantee in new Theorem~\ref{thm:stats}. The MLE error matches with statistical lower bound and ensures the dTBM achievability above the statistical critical value. We have also revised the exposure of Theorems~\ref{thm:stats}-\ref{thm:comp} to include both impossibility and achievability results. See our new Section~\ref{sec:statlimit} and Section~\ref{sec:complimit}. See our response to {\bf \# 2 Major Comments from Editor} for details.

    \item Regarding ``The stated upper bound assume some uniformity in $\theta$". We have added a new Section~\ref{sec:prelim} to discuss the rational for the balanced degree assumption~\eqref{eq:degree}. Briefly, we focus on the regime with balanced heterogeneity due to the relationship between dTBM performance and core tensor signal $\Delta_{\min}^2$. Indeed, without the balance assumption, the phase transition of dTBM can be arbitrarily worse than that of TBM. We quote our new discussion below:

New Section~\ref{sec:prelim}:
    \begin{quote}
    ``...In our work, we focus on the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
\begin{equation}\label{eq:degree}
{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
\end{equation}
The following lemma provides the rational of balanced degree assumption. We show the close relation between angle gaps in the mean tensor $\tX$ and the core tensor $\tS$ under balanced degree heterogeneity.

\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}. Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}. Then, for all $i,j$ such that $z(i) \neq z(j)$, we have
\begin{equation}
 % \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  },
 \cos(\mX_{i:}, \mX_{j:})\asymp  \cos(\mS_{z(i):}, \mS_{z(j):}),
\end{equation}

where $\mX =\mat(\tX)$ and $\mS = \mat (\tS)$.
\end{lem}

%Intuitively, the performance of dTBM is determined by both the core tensor and degree heterogeneity. Since we view degrees as nuisance parameters, our main interest is to characterize dTBM performance with respect to the signal $\Delta^2_{\min}$ in the core tensor. By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}. Therefore, the balance assumption helps to exclude the cases in which the degree heterogeneity distorts the dTBM performance.

In practice, an estimation algorithm has access to a noisy version of $\tX$ but not $\tS$. Our goal is to establish the algorithm performance with respect to the signal $\Delta^2_{\min}$ in the core tensor. By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}. Therefore, the balanced degree assumption helps to exclude the cases in which the degree heterogeneity distorts the algorithm garantees.

%Indeed, as we will see in Sections~\ref{sec:statlimit}-\ref{sec:complimit}, under the balance assumption~\eqref{eq:degree}, the $\Delta_{\min}^2$ is sufficient to establish the dTBM phase transition.

%In contrast, a highly-skewed degree violating~\eqref{eq:degree} may lead to a worse performance of dTBM than that of TBM. 
Here, we provide an example to illustrate the insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees.

% The balance assumption \eqref{eq:degree} ensures the sufficiency of core tensor signal $\Delta_{\min}^2$ to describe the phase transition of dTBM.
% Without balanced degree,

%The dTBM phase transition without balanced degree may be arbitrary worse than that with balanced degree.

\begin{example}[Insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees] Consider an order-2 $(p,p)$-dimensional dTBM with core matrix
\begin{equation}\label{eq:modelexample}
    \mS = \begin{pmatrix} 1 &a\\
    1 & -a
    \end{pmatrix}, \quad \text{and} \ \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = p^m \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
\end{equation}
where $m\geq 0$ is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the angle gap of the mean tensor, defined by
\begin{equation}\label{eq:delta_x}
    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \mX = \mat(\tX). %\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
\end{equation}
Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
\begin{equation}
    \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2} \asymp p^{-1/2}, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-1/2-m}.
\end{equation}
%Take $ a = p^{-1/4}. We have 
%\begin{equation}
   % \Delta_{\min}^2 = p^{-1/2} \quad \text{and} \quad    \Delta_{\mX}^2 = {1\over p^{m-1/2}+1} = \min(p^{1/2-m},1). 
    %{2 a^2 \over M+a^2} = p^{-1-\epsilon}.
%\end{equation}
% \begin{align}
%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 =  \Delta_{\mX}^2 = p^{-1}; \\
%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \Delta_{\mX}^2 = p^{-1}.
% \end{align}
%By the following Theorem~\ref{thm:stats}, no estimator achieves exact recovery of dTBM under $\Delta_{\mX}^2 < p^{-1}$. 
%Under the balanced degree assumption, the dTBM is computationally solvable based on our Theorem~\ref{thm:comp} in Section~\ref{sec:complimit}.
Based on our theory in later Sections, the stat-computational achievability of dTBM occurs when $\Delta^2_{\mX}>p^{-1}$; that is, the dTBM estimation depends on the relative magnitude of $m$ vs. $1/2$. In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM. %However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
%$-1\leq \varepsilon \leq 1/2$.
 %($\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}). 
\end{example}

\begin{rmk}[Flexibility in balanced degree assumption] One important note is that our balance assumption~\eqref{eq:degree} does not preclude the mild degree heterogeneity. In fact, within each of the clusters, we allow the highest degree at the order $\tO(p)$, whereas the lowest degree at the order $\Omega(1)$. This range is more relaxed than previous work \citep{gao2018community} that restricts the highest degree in the sub-linear regime $o(p)$ and the lowest degree at the order $\Omega(1)$. 
\end{rmk}

\begin{rmk}[Similar assumptions in literature]
%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
Similar degree regulations are not rare in literature. In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure balancedness across communities. %Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees. " 
In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. 
\end{rmk}

    \end{quote}

\end{enumerate}

%Detailed revisions in Section~\ref{sec:limits} are in our response to \textbf{\# 2 Major Comments from Editor}.

% For the statistical lower bound, we have revised the proof of minimax bound with general degree heterogeneity and the lower bound would not significantly increases. We have added the MLE guarantee to ensure the achievability above the statistical critical value. We have also revised the exposure of Theorem~\ref{thm:stats} and Theorem~\ref{thm:comp} to include both impossibility and achievability results.

% We have changed the exposure of Theorems~\ref{thm:stats} and \ref{thm:comp} to include both the impossibility and achievement of dTBM problem when signal is below and above the critical values. We quote the revised paragraphs in Sections~\ref{sec:statlimit} and \ref{sec:complimit}.

%We focus on the regime with balanced $\mtheta$ due to our main interest on the relationship between dTBM performance and core tensor signal $\Delta_{\min}^2$. Indeed, without balance assumption, the phase transition of dTBM can be arbitrarily worse than that with balanced assumption or that of TBM. We have added a new Section~\ref{sec:prelim} to discuss the settings. \fixme{Miaoyan}{Quote your discussion here.}
% For the balance assumption on degree heterogeneity~\eqref{eq:degree}, we have added a preliminary section~\ref{sec:prelim} in Section~\ref{sec:limits} to discuss our signal notion and the balance regulation on $\mtheta$. The balance assumption is proposed to describe the performance of dTBM by the signal only in core tensor. We show an example to illustrate that the phase transition of dTBM would be arbitrary worse than TBM with unbalanced heterogeneity. Our statistical and computational limits are sharp under the regime of balance degree. We quote the discussion in the preliminary Section~\ref{sec:prelim} here:

% Our statistical lower bound is derived under the dTBM since we consider the estimator space without the equal degree assumption. The condition (i), $\theta(i) = 1$ for all $i \in [p]$, was assumed on true heterogeneity for technical convenience. We have now relaxed the equal degree condition (i) in the proof of Theorem 2 to avoid the ambiguity. We quote the revised proof below:

% \begin{quote}
%     ``
%     ... Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following two properties:
% \begin{equation}\label{eq:construction}
% \text{(i)}\ 
%     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(ii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
% \end{equation}
% for all $k \in [K]$.

% Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. The construction of $T$ is precisely the same as \citet[Proof of Theorem 2]{gao2018community}. By the constraint $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)|, a \in [r]$ in parameter space (3), the $\max_{i \in T^c} \theta(i) \leq C$ for some positive constant $C$. Then, following the proof of~\citet[Theorem 2]{gao2018community}, ...."
% \end{quote}

% We have showed that MLE matches the statistical limit under the balance assumption in a new theorem. See our response to \textbf{\# 2 Major Comments from Editor.}

% \fixme{Jiaxin}{Need to rethink this question about the signal notion and phase transition.}
% {
% \color{red}

% The phase transition for dTBM is arbitrary worse than that of TBM, no matter with angle-based or Euclidean-based SNR. See Example 1. That's why we need to propose angle based SNR. 
% %To see the necessity of balance, we need to show the SNR phase transition will arbitrary worse than that with balance assumption.

% Notice that 
% \begin{equation}
%     \{\text{$\mtheta$ is balanced} ,  \Delta_{\mS, \min}^2 \coloneqq \min \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}^2 \leq p^{\gamma} \} \quad \subset \quad \{ \Delta_{\mA, \min}^2 \coloneqq \min \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}^2 \leq  p^{\gamma}\},
% \end{equation}
% where $\mA = \mS (\mM^T \mTheta^T)^{\otimes K-1}$.

% Our current statistical and computational limits are sharp under the former notion, $\Delta_{\mS, \min}^2$ with balance. The statistical lower bound and initialization guarantee still hold with latter notion $\Delta_{\mA, \min}^2$; but the computational limits, algorithm guarantees, MLE are not easy to extend with $\Delta_{\mA, \min}^2$.

% Specifically, in the computational lower bound,  we pick up a special example with $\theta = 1$. This special case lies in the parameter spaces generated by both signal notions. We apply Han's result to the special case, and this indicates computational lower bound can be much higher than current one; i.e., when $\Delta_{\mS, \min}^2$ with balance or $\Delta_{\mA, \min}^2 $ is larger than $p^{-K/2}$, it still may be computationally inefficient. The good news is: when $\Delta_{\mS, \min}^2 = p^{-K/2} \log p$ with balance, we have a polynomial time algorithm achieves exact recovery. Therefore, the threshold  $p^{-K/2}$ is sharp if we use the signal notion $\Delta_{\mS, \min}^2$ with balance.

%  In the iteration, we do not know the $\mtheta$ in practice, do not estimate $\tX$ in algorithm, and need the gap between folded vectors $\mS_{a:}$'s. Hence, it may not be easy to show the algorithm guarantee with later notion $\Delta_{\mA, \min}^2$; the MLE upper bound relies on iteration results  may also not be easy to extend.

% }

\item \textit{Please provide some intuition on Deﬁnition 2, particularly Eqn (10).}

\textbf{Response:} We have added more explanations for Definition~\ref{def:stable} and equation~\eqref{eq:local} in Section~\ref{subsec:angle}:
\begin{quote}
    ``... Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_\theta(\bar z)$ represents the relative cluster sizes weighted by degrees. ...The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap $\Delta_{\min}$. Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled. The stability assumption is proposed for technical convenience, and we relax this condition in numerical studies; see Section~\ref{sec:simulation}."
\end{quote}

%We thank reviewer for the comment. Previous Definition 2 and equation (10) ensures the small gap between raw cluster size $|\hat z^{-1}(b)|$ and weighted cluster size $|\sum_{i \in \hat z^{-1}(b)} \theta(i)|$. With a better technique, we have shown the Definition 2 and the local linear stability condition on the heterogeneity $\mtheta$ are unnecessary for the algorithm guarantee. We have removed related parts from our manuscript.
\end{enumerate}
